SHARP - II. Beware the effect of edge-on discs when interpreting the anomalous flux-ratios of gravitationally lensed quasars

### 2015-09-08 papers/B1555_EdgeOn/B1555_letter.tex
shrink the size of figures 84mm->72mm
take out chi^2 in text

We use the lens modelling code {\sc gravlens} \citep{Kee01} to model the compact radio components. The inputs to the model are the observed image positions measured from the MERLIN radio observations of \citet{Marlow99} as all of the images are significantly detected in these data. The flux-ratio measurements are taken from \citet{K03}, who used half a year of MERLIN monitoring to obtain flux-ratio curves. In total, there are 11 constraints to the mass model provided by the observational data. As a first trial of modelling, we re-create the singular isothermal ellipsoid without external shear (SIE; 7 free parameters) model of \citet{Marlow99} to check the performance of a simple lensing potential. Consistent with previous studies \citep{Marlow99,Xu15}, a single elliptical potential model cannot reproduce the observed flux-ratios in B1555+375 (see Table~\ref{tab:results}; reduced $\chi^2 =97.3$).

@@

The best-fit model parameters and 68\% CL errors are listed in Table~\ref{tab:model} and the lens model is illustrated in Fig.~\ref{fig:model}. Table \ref{tab:results} also shows the comparison between the observed radio data and the model predicted image positions and flux ratios. We find that by adding an elongated exponential disc to a smooth potential model there is significant improvement in the predicted flux-ratios when compared to the results for the SIE-only model, with a reduced $\chi^2 = 2.7$. The positions are all in agreement, and the flux-ratios are consistent at the $2\sigma$ level, with excellent agreement now obtained for the two merging images A and B. The differences in fluxes may due to the source could be slightly extended sphere while our model uses point-source. The mass model requires that the exponential disc component makes up about 10 per cent of the total mass within the Einstein radius (1.7~kpc for a lens redshift of $z_l = 1$), which is not unreasonable. We note that there is a significant 0.036~arcsec offset between the SIE and exponential disc components, which is likely due to there being an additional ellipticity in the model (e.g. from an external shear) that is being partly compensated for by the disc component. Again, more complex mass models that use the full extent of the gravitational arc and the compact lensed images from future high resolution radio imaging will be able to confirm this.

\begin{figure}
%\includegraphics[width=84mm]{1555_ao_merlin_overlay.eps}
\includegraphics[width=84mm]{fig2_twocontour_g.eps}
\caption{High-resolution $K^\prime$-band imaging of B1555+375 with contours from the MERLIN observations (red) of \citet{Marlow99} and the VLBA observations presented here overlaid.}
\label{fig:merlin}
\end{figure}

\begin{figure}
\includegraphics[width=84mm]{gravlens_exp_try5_plot.eps}
\caption{Observed radio (blue pluses) and model-predicted (red open circles) image positions of B1555+375. The lens-plane critical curves are shown with the blue solid line and the source-plane caustics are shown with the red dash line. The position of the source is at ($-0.1953$, $-0.1494$), which is marked by the red filled circle. The centroid positions of the two mass components of the lens (an SIE and an exponential disc) are marked by the black filled triangles.}
\label{fig:model}
\end{figure}

>>>>

We use the lens modelling code {\sc gravlens} \citep{Kee01} to model the compact radio components. The inputs to the model are the observed image positions measured from the MERLIN radio observations of \citet{Marlow99} as all of the images are significantly detected in these data. The flux-ratio measurements are taken from \citet{K03}, who used half a year of MERLIN monitoring to obtain flux-ratio curves. In total, there are 11 constraints to the mass model provided by the observational data. As a first trial of modelling, we re-create the singular isothermal ellipsoid without external shear (SIE; 7 free parameters) model of \citet{Marlow99} to check the performance of a simple lensing potential. Consistent with previous studies \citep{Marlow99,Xu15}, a single elliptical potential model cannot reproduce the observed flux-ratios in B1555+375 (see Table~\ref{tab:results}). 
%; reduced $\chi^2 =97.3$

@@

The best-fit model parameters and 68\% CL errors are listed in Table~\ref{tab:model} and the lens model is illustrated in Fig.~\ref{fig:model}. Table \ref{tab:results} also shows the comparison between the observed radio data and the model predicted image positions and flux ratios. We find that by adding an elongated exponential disc to a smooth potential model there is significant improvement in the predicted flux-ratios when compared to the results for the SIE-only model. 
% with a reduced $\chi^2 = 2.7$
The positions are all in agreement, and the flux-ratios are consistent at the $2\sigma$ level, with excellent agreement now obtained for the two merging images A and B. The differences in fluxes may due to the source could be slightly extended sphere while our model uses point-source. The mass model requires that the exponential disc component makes up about 10 per cent of the total mass within the Einstein radius (1.7~kpc for a lens redshift of $z_l = 1$), which is not unreasonable. We note that there is a significant 0.036~arcsec offset between the SIE and exponential disc components, which is likely due to there being an additional ellipticity in the model (e.g. from an external shear) that is being partly compensated for by the disc component. Again, more complex mass models that use the full extent of the gravitational arc and the compact lensed images from future high resolution radio imaging will be able to confirm this.

\begin{figure}
%\includegraphics[width=84mm]{1555_ao_merlin_overlay.eps}
\includegraphics[width=72mm]{fig2_twocontour_g.eps}
\caption{High-resolution $K^\prime$-band imaging of B1555+375 with contours from the MERLIN observations (red) of \citet{Marlow99} and the VLBA observations presented here overlaid.}
\label{fig:merlin}
\end{figure}

\begin{figure}
\includegraphics[width=72mm]{gravlens_exp_try5_plot.eps}
\caption{Observed radio (blue pluses) and model-predicted (red open circles) image positions of B1555+375. The lens-plane critical curves are shown with the blue solid line and the source-plane caustics are shown with the red dash line. The position of the source is at ($-0.1953$, $-0.1494$), which is marked by the red filled circle. The centroid positions of the two mass components of the lens (an SIE and an exponential disc) are marked by the black filled triangles.}
\label{fig:model}
\end{figure}
